Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2014-02-10 paper/paper.tex
put combo plot (not yet fixed up) into paper Tested everything on bennet.
added lower threshhold for mannik shape to paper Tested everything on node08.

In order to explain this theoretically, we have performed stability
analysis of the five differential equations in an infinite slab that
shows a characteristic half wavelength in perturbative spatial
oscillations at which the system becomes unstable.  This length is
dependent upon slab thickness as shown in
Figure~\ref{fig:stability-analysis}.  When increasing the lengths and
widths of our simulated flattened cells, the steady, two pole
oscillations seem to break down when the the shortest distance across
the cell is approximately equal to the characteristic half wavelength
at the particular thickness at which we simulate (our simulations of
cells with thickness $0.25\mu m$ corresponds to half wavelengths of
$2.13\mu m$).  Huang~\cite{huang2003dynamic} performs a similar
analysis for a cylindrical shape and finds characteristic half
wavelengths of $2\mu m$.

@@

@@

Looking at at few periods in detail.  We find that there is a small
increase in cytoplasmic MinE always directly before the the beggining
of climb in the membrane bound MinE peak.  Also, there is at any one
time a very large amount more of minE that is membrane bound then
cytoplasmic in any ones section.  A difference of a factor of roughly
20 when membrane MinE is at its peak, and roughly 12 when membrane
MinE is at it's minimum.  There seem to be two modes of oscillation,
each exhibiting a density peak maxima that moves back and forth
diagonaly from one corner to the opposite, as if there were two
overlapping, criss-crossed cells oscillating.  In the box plot this is
seen as roughly coinciding membrane minD peaks in the lower left and
right corners, followed one pi later in peaks in the upper right and
upper left corners.  Interesting that this did not exhibit
counterclockwise motion that Huangs 2008 paper might suggest.  Should
run sims that start in the wacky way that just started running with
triangles.
\newline

>>>>

In order to explain this theoretically, we have performed stability
analysis of the five differential equations in an infinite slab that
shows a characteristic half wavelength in perturbative spatial
oscillations at which the system becomes unstable.  This length is
dependent upon slab thickness as shown in
Figure~\ref{fig:stability-analysis}.  When increasing the lengths and
widths of our simulated flattened cells, the steady, two pole
oscillations seem to break down when the shortest distance across the
cell is approximately equal to the characteristic half wavelength at
the particular thickness at which we simulate (our simulations of
cells with thickness $0.25\mu m$ corresponds to half wavelengths of
$2.13\mu m$).  Huang~\cite{huang2003dynamic} performs a similar
analysis for a cylindrical shape \fixme{is it cylindrical or 1d?} and
finds characteristic half wavelengths of $2\mu m$.

@@

\begin{figure}
  \includegraphics[width=3cm]{../data/shape-randst/plots/box-plot_D--randst-25-800-800-9500-1500}
  \includegraphics[width=3cm]{../data/shape-randst/plots/box-plot_D--randst-25-1000-1700-9400-1500}
  \caption{Box plots of Mannik shapes for sizes at the lower stability
    threshhold.  Both shapes have longer axis of roughly 2.2$\mu m$.}
  \label{box-mannik}
\end{figure}
Decreasing the cell size in succesive steps, we can see as well the
critical size at which the cell oscillations are quickly damped out.
Figure~\ref{box-mannik} shows box plots of the largest cells simulated
whose oscilltions relax into the steady state solution.  Both have a
long axis of roughly 2.2$\mu m$, in agreement with theory.

@@

Looking at at few periods in detail.  We find that there is a small
increase in cytoplasmic MinE always directly before the beggining
of climb in the membrane bound MinE peak.  Also, there is at any one
time a very large amount more of minE that is membrane bound then
cytoplasmic in any ones section.  A difference of a factor of roughly
20 when membrane MinE is at its peak, and roughly 12 when membrane
MinE is at it's minimum.  There seem to be two modes of oscillation,
each exhibiting a density peak maxima that moves back and forth
diagonaly from one corner to the opposite, as if there were two
overlapping, criss-crossed cells oscillating.  In the box plot this is
seen as roughly coinciding membrane minD peaks in the lower left and
right corners, followed one pi later in peaks in the upper right and
upper left corners.  Interesting that this did not exhibit
counterclockwise motion that Huangs 2008 paper might suggest.  Should
run sims that start in the wacky way that just started running with
triangles.
\newline
